Building Markov State Models for Periodically Driven Non-Equilibrium Systems

### 2014-11-05 writedown/period.noneq.msm/paper.tex
final revise from CS

Biomolecular systems under non-equilibrium conditions caused by external fields, especially
systems under periodic forcing, have attracted increasing
interest recently.  For example, the potential effects of electromagnetic
radiation on the human body tissue (e.g.~DNA, protein, and membrane)  has been extensively
investigated in a vast number of articles, with the following list just representing an incomplete selection~\cite{bohr2000microwave, bohr2000microwave-1, dePomerai2000cell,
  dePomerai2003microwave, mancinelli2004non, inskip2001cellular, bekard2013electric, budi2005electric,
  budi2007effect, budi2008comparative, astrakas2012structural,
  damm2012can, english2009nonequilibrium, solomentsev2012effects}.  Molecular dynamics (MD) simulations have proved particularly useful for understanding
the response of biomolecular conformations to external fields because of their ability to  resolve molecular
details that \recheck{sometimes} cannot be resolved in experiments. Only
recently, a \recheck{dynamical} non-equilibrium MD simulation (D-NEMD) method with local thermostating has been proposed~\cite{wang2014exploring} that allows for studying 
non-equilibrium processes in a statistically reliable and thermodynamically consistent way. 
Despite the significance of the non-equilibrium phenomena, the
analysis of the non-equilibrium MD simulations mainly follows standard approaches, and reliable tools for quantitative description of
the essential conformational dynamics of the molecular system under external forcing are still lacking.

@@

The outline of the article is as follows: 
In Sec.~\ref{sec:disc}, we discuss the temporal and spatial
discretizations needed to construct an MSM. We consider spatial discretization
of the dihedral angle space in the traditional sense of full partition MSMs~\cite{pande2010everything,A19-29}.
In the temporal direction, the non-equilibrium process
is discretized utilizing Floquet's theorem. This results in a time-homogeneous, 
but not necessarily reversible, that is, irreversible Markov process. Since  full spatial partition does not make any sense
for high-dimensional systems, we next show how to construct a \emph{few-state} MSM based on milestoning~\cite{schuette2011markov,A19-29} of the discretized irreversible Markov process in Sec.~\ref{sec:build-msm}.
The validity of the discretizations and the resulting MSM is checked in Sec.~\ref{sec:alanine}
by comparing the kinetic fingerprints given by the MSM to brute force
non-equilibrium \recheck{MD} simulations of alanine dipeptide under oscillating EF.
The findings are summarized in Sec.~\ref{sec:conclusion} including a chart presenting
the workflow of MSM building for non-equilibrium systems, and a list of open questions.

@@

\section{Non-equilibrium molecular dynamics and its discretization}
\label{sec:disc}
\recheck{We consider diffusive molecular dynamics} in an energy landscape $V$ driven by the time-dependent external driving force $E(t)D(x_t)$ with the
$T$-periodic external field $E(t)$:
\begin{align}
  \label{eq:disc-1}
  d\vect x_t = \Big(-\nabla V(\vect x_t) + E(t) D(\vect x_t)\Big)dt + \sqrt{2\beta^{-1}} d\vect w_t, 
\end{align}
where $x_t\in\Omega$ denotes the state of the molecular system at time $t$ in state space $\Omega$, $\vect w_t$ denotes standard $n$-dimensional Brownian motion, and $\beta$ the inverse temperature,
i.e.~$\beta = 1/(k_B\mathcal T)$.
\recheck{The Hamiltonain or Langevin dynamics can be treated in the same way as that will be presented, so for sake of simplicity we focus on the discussion of diffusive dynamics.}
The propagation of probability
densities $\rho=\rho(\vect x,t)$ based on this kind of dynamics in the sense
of $\rho(\vect x,t)dx=\prob[\vect x_t\in [\vect x,\vect x+d\vect x)]$ is governed by
Fokker-Planck equation:
\begin{align}
  \label{eq:disc-fp}
  \frac{\partial \rho}{\partial t} = \ml^\dagger(t) \rho,
\end{align}
where $\ml^\dagger(t)$ is the adjoint of the generator
\begin{align}
  \label{eq:disc-3}
  \ml(t)=\beta^{-1}\Delta_{\vect x}+\Big(-\nabla_x V(\vect x) + E(t)D(\vect x)\Big)\cdot\nabla_{\vect x},
\end{align}
where $\Delta_{\vect x}$ denotes the Laplacian operator and $\nabla_{\vect x}$
the nabla-operator wrt to $\vect x$. 
The periodicity of the external driving force induces the periodicity of the generator,
i.e.~$\ml(t) = \ml(t+T)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial discretization: Master equation}
We will now introduce an appropriate spatial discretization of this kind of non-equilibrium MD -- \recheck{\bluec{(why saying this red sentence? we need the spatial discretization for calculating the core sets...)} this is done for reasons of simplicity only; we could completely avoid it for the price of more technical arguments.}
For achieving this discretization, we 
introduce a partition of state space $\Omega$ into a finite number of disjoint
sets $\{ \Omega_1, \cdots, \Omega_N\}$ satisfying $\Omega = \cup_i \Omega_i$,
$\Omega_j\cap \Omega_j = \emptyset,\ \forall i\neq j$.
Utilizing the procedure described in Ref.~\cite{latorre2011structure} the original Fokker-Planck equation~(\ref{eq:disc-fp})
is discretized, resulting in a time-inhomogeneous Markov jump process in state
space $S = \{1, \cdots, N\}$ with time-dependent rate
matrix $\vect L(t) \in \mathbb R^{N\times N}$ satisfying
\begin{align}\label{eq:disc-4}
\sum\limits_{j=1}^N L_{ij}(t) & =  0\\ \label{eq:disc-5}
L_{ij}(t) & \ge  0, \quad i\not= j\\
L_{ij}(t) & =  L_{ij}(t+T)
\end{align}
for all real time $t\geq 0$.
Moreover, the rate matrix $L$ has the form $\vect L(t)=\vect L_0+E(t)\vect L_1$
where $E(t)$ is periodic with period $T>0$.
In analogy to \eqref{eq:disc-fp}, the Markov jump process generated by
$\vect L(t)$ transports probability distributions according to the associated Master equation
\begin{align}
  \label{eq:disc-master}
  \frac{d\vect p(t)}{dt} = \vect L^{\top}(t)\cdot \vect p(t)
\end{align}
where $\vect L^{\top}(t)$ denotes the matrix transpose of $\vect L(t)$, $\vect p(t)$ is an $N$-vector denoting the probability distribution on $S$ at time $t$, $p(i,t)$, for example, the probability to be in state $i$ (which corresponds to set $\Omega_i$) at time $t$.
As usual the properties (\ref{eq:disc-4}) and (\ref{eq:disc-5}) of
$\vect L(t)$ guarantee that the total probability mass is conserved,
i.e., if $p(i,0)\ge 0$ componentwise, then $p(i,t)\ge 0$ and $\sum_i
p(i,t) = \sum_ip(i,0)$.
% The solution of the master equation need no be
% periodic.
The temporal evolution of the probability distribution $\vect p(t)$ can be formally written
\begin{align}  \label{eq:disc-8}
\vect p(t)=\myphi(t)\vect p(0)
\end{align}
by using the
associated propagator matrix $\myphi(t)\in\mathbb R^{N\times N}$ that solves
\begin{align}
  \label{eq:disc-master-phi}
  \frac{d}{dt}\myphi(t) = \vect L^{\top}(t)\myphi(t), \quad \myphi(0) = \id .
\end{align}
Since the last equation can be considered column-wise, the propagator matrix inherits column-wise conservation properties:
$\Phi_{ij}(t) \ge  0$
and $\sum\limits_{i=1}^N \Phi_{ij}(t)  =  1$,
that is, $\myphi^{\top}(t)$ is a
stochastic matrix satisfying $\myphi^{\top}(t)\vect e=\vect e$
with $\vect e=(1,\ldots,1)^{\top}\in \mathbb R^N$.
Regarding these considerations, we find
\begin{align}
\label{eq:disc-10}  
\myphi_{ij}(t)=\prob\left(\vect X_t=i\mid \vect X_0=j \right),
\end{align}
where $\vect X_t$ denotes the Markov process generated by $\vect L(t)$.

@@

Since $P$ is a stochastic matrix, its eigenvalues are contained in the unit circle in
the complex plane, i.e., each eigenvalue $\lambda$ (potentially complex-valued)
satisfies $|\lambda|\le 1$. Furthermore $\lambda=1$ is an
eigenvalue with right eigenvector $\vect e=(1,\ldots,1)^{\top}$ and a left eigenvector $\mymu$
satisfying
$\mymu^{\top}P=\mymu^{\top}$.
From now on, we assume $P$ to be irreducible and aperiodic such that the Perron-Frobenius theorem \recheck{holds, so the eigenvector corresponding to} the eigenvalue $\lambda=1$ is  non-negative componentwise, and unique (up to normalization $\sum_i\mu(i)=1$). In this case $\mymu$ is the stationary measure in the sense that
$
 \mymu^{\top} P^m = \mymu^{\top},\ m\in\mathbb N
$,
% \begin{align}
% \label{eq:floq-14}  
% \end{align}
and (more precisely) the asymptotic evolution of an initial probability distribution $\vect p(t=0)$ by the process satisfies
$
\vect p^{\top}(0)\,P^m \to \mymu^{\top},\ m\to\infty,
$
% \begin{align}
% \label{eq:floq-15}  
% \end{align}
so that $\mymu$ can be seen as the quasi-stationary distribution of the non-stationary process.

@@

To illustrate how the discretization works in practice we take the alanine dipeptide system under an oscillatory EF,
as an example, the NEMD simulation of which was
extensively studied in Ref.~\cite{wang2014exploring}.
The system was simulated in a $2.7\times 2.7\times 2.7$~$\textrm{nm}^3$ periodic simulation
region, with one alanine dipeptide molecules described by the CHARMM27 force field~\cite{foloppe2000all} dissolved in 641 TIP3P water molecules~\cite{jorgensen1983comparison}.
The grid-based energy correction map (CMAP)~\cite{mackerell2004extending} was used
to correct the backbone dihedral angle energies.
All simulations were performed by a home-modified
Gromacs~4.6.5~\cite{pronk2013gromacs} with CHARMM27 force field implemented~\cite{bjelkmar2010implementation}.
The alanine dipeptide was put into the local thermostating
environment, with a spherical dynamical region of radius~1.0~nm
centered at the alpha-carbon.
The Langevin thermostat with target temperature $\mathcal T = 300$~K
and timescale $\tau_T = 0.1$~ps was
coupled to the thermostated region.
The whole system was coupled to the Parrinello-Rahman barostat~\cite{parrinello1981polymorphic} (in standard Gromacs implementation) with $\tau_P = 2.0$~ps to
keep the system at 1~Bar. The non-equilibrium trajectories
were integrated by the Leap-frog scheme with a time-step of 0.002~ps.
The short-range van der Waals interactions were cut-off at 1.00~nm, and were smoothed from
0.95~nm to 1.00~nm by the ``\texttt{shift}'' method provide by Gromacs.
The energy conserving 
Particle Mesh Ewald (PME)~\cite{darden1993pme, essmann1995spm} method (``\texttt{pme-switch}'') was
used to compute the long-range electrostatic interaction,
with the same real-space cut-off radius as the van der Waals interactions.
\recheck{The Gromacs default Fourier spacing of 0.12~nm and B-spline interpolation order of 4 were adopted.}
The splitting
parameter was optimized with respect to the electrostatic force computing
accuracy by Gromacs tool \texttt{g\char`_pme\char`_error}~\cite{wang2010optimizing}.
The neighbor list was updated every 5 time-steps with a list-building radius 1.20~nm.
All hydrogen involving covalent bonds were constrained by the LINCS algorithm~\cite{hess1997lincs}, except the water molecules that were constrained by the SETTLE algorithm~\cite{miyamoto2004settle}.
The whole system was driven by a periodic electric field
$E(t) = E_0\sin(2\pi t/T)$ and $D(x) = (1,0,0)^{\top}$
with intensity of the field being $E_0 = 1.0$~V/nm and period being
$T = 10$~ps.
The 20,000 branching trajectories were simulated from 20,000
initial configurations that sample the equilibrium distribution.
% \redc{Write a lot of details on NEMD.}
The equilibrium configurations were prepared by an equilibrium MD simulation
of length $10^6$~ps, along which snapshots were saved every 50~ps.
% In the
% equilibrium simulation the system was coupled to a global Langevin thermostat
% with coupling timescale $\tau_T = 0.5$~ps.
The branching NEMD
trajectories were each 4,000~ps long, and the system reached
non-equilibrium quasi-stationary state in roughly 300~ps.

For this periodically driven molecular system we will first show how to choose an appropriately fine \recheck{spatial} discretization. After validating this discretization we will consider
the \recheck{temporal} discretized dynamics generated by the Floquet transition matrix $\vect P$
in comparison to the original NEMD simulation. Finally we will coarse grain this description further by construction of a $3$ state Markov State Model that is able to describe the long-term kinetics of the system correctly.

@@

An important check is the consistency between the
stationary probability density of $\Phi(T)$ (i.e.~the leading eigenvector $\mymu$) and that 
estimated form the original NEMD simulation,
% along which only integral periods $mT$ are considered:
\begin{align}
  \label{eq:num-tmp1}
  \rho_{\textrm{st}}(\phi,\psi) = \lim_{m\rightarrow\infty} \rho (\phi,\psi,mT),
\end{align}
On each NEMD branching trajectory the initial 320~ps are discarded and
the rest of the trajectory in time interval $[320,4000]$~ps is \recheck{averaged} to estimate
the quasi-stationary probability distribution $\rho_{\textrm{st}}$. 
$P$ is computed as described above, and then $\mymu$ is computed as its leading eigenvector. In order to make it comparable to the free energy in the equilibrium case, we take
the logarithm of the distributions, i.e.~$F_{\textrm{st}}(\phi,\psi)=
-k_B\mathcal T\log \rho_{\textrm{st}}(\phi,\psi)$
for NEMD and $F_{\textrm{st}}(\phi,\psi)=
-k_B\mathcal T\log \mu(\phi,\psi)$ for $P$, where $k_B$ is the
Boltzmann constant and $\mathcal T$ is the temperature of the system.
% Under the aformentioned discretization, the stationary distribution is
% defined for the each bin of the dihedral angle space, i.e.~$F_{pq} =
% \int_{ph}^{(p+1)h}d\phi\int_{qh}^{(q+1)h}d\psi
% F_{\textrm{st}}(\phi,\psi), \ 0\leq p,q<N_\dih-1$, where $h = 360/N_\dih$ being
% the size of the bin, $p$ and $q$ here being the bin indexes.
The
results are  compared in Fig.~\ref{fig:num-1}. A good consistency between
the  NEMD simulation and 
$P$ is observed.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Forward and backward committors}
\label{sec:alanine-committor}
\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-msm-1.eps}\\
  \vskip -.7cm
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-msm-1.eps}\\
  \vskip -.7cm
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-msm-1.eps}
  \vskip -.5cm
  \caption{\recheck{The forward $q^\fwd_{\confaa}$ and backward 
    $q^\bwd_{\confaa}$ committors
    and their difference $q^\fwd_{\confaa} - q^\bwd_{\confaa}$
    computed by the NEMD simulations (left
    column) and the  discretized dynamics (right column).}}
  \label{fig:num-3}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-msm-2.eps}\\
  \vskip -.7cm
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-msm-2.eps}\\
  \vskip -.7cm
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-msm-2.eps}\\
  \vskip -.5cm
  \caption{\recheck{The forward $q^\fwd_{\beta}$ and backward
    $q^\bwd_{\beta}$ committors
    and their difference  $q^\fwd_{\beta} - q^\bwd_{\beta}$ 
    computed by the NEMD simulations (left
    column) and the  discretized dynamics (right column).}}
  \label{fig:num-4}
\end{figure}

>>>>

Biomolecular systems under non-equilibrium conditions caused by external fields, especially
systems under periodic forcing, have attracted increasing
interest recently.  For example, the potential effects of electromagnetic
radiation on the human body tissue (e.g.~DNA, protein, and membrane)  has been extensively
investigated in a vast number of articles, with the following list just representing an incomplete selection~\cite{bohr2000microwave, bohr2000microwave-1, dePomerai2000cell,
  dePomerai2003microwave, mancinelli2004non, inskip2001cellular, bekard2013electric, budi2005electric,
  budi2007effect, budi2008comparative, astrakas2012structural,
  damm2012can, english2009nonequilibrium, solomentsev2012effects}.  Molecular dynamics (MD) simulations have proved particularly useful for understanding
the response of biomolecular conformations to external fields because of their ability to  resolve molecular
details that sometimes cannot be resolved in experiments. Only
recently, a non-equilibrium MD simulation (D-NEMD) method with local thermostating has been proposed~\cite{wang2014exploring} that allows for studying 
non-equilibrium processes in a statistically reliable and thermodynamically consistent way. 
Despite the significance of the non-equilibrium phenomena, the
analysis of the non-equilibrium MD simulations mainly follows standard approaches, and reliable tools for quantitative description of
the essential conformational dynamics of the molecular system under external forcing are still lacking.

@@

The outline of the article is as follows: 
In Sec.~\ref{sec:disc}, we discuss the temporal and spatial
discretizations needed to construct an MSM. We consider spatial discretization
of the dihedral angle space in the traditional sense of full partition MSMs~\cite{pande2010everything,A19-29}.
In the temporal direction, the non-equilibrium process
is discretized utilizing Floquet's theorem. This results in a time-homogeneous, 
but not necessarily reversible, that is, irreversible Markov process. Since  full spatial partition does not make any sense
for high-dimensional systems, we next show how to construct a \emph{few-state} MSM based on milestoning~\cite{schuette2011markov,A19-29} of the discretized irreversible Markov process in Sec.~\ref{sec:build-msm}.
The validity of the discretizations and the resulting MSM is checked in Sec.~\ref{sec:alanine}
by comparing the kinetic fingerprints given by the MSM to brute force
non-equilibrium MD simulations of alanine dipeptide under oscillating EF.
The findings are summarized in Sec.~\ref{sec:conclusion} including a chart presenting
the workflow of MSM building for non-equilibrium systems, and a list of open questions.

@@

\section{Non-equilibrium molecular dynamics and its discretization}
\label{sec:disc}
We consider diffusive molecular dynamics in an energy landscape $V$ driven by the time-dependent external driving force $E(t)D(x_t)$ with the
$T$-periodic external field $E(t)$:
\begin{align}
  \label{eq:disc-1}
  d\vect x_t = \Big(-\nabla V(\vect x_t) + E(t) D(\vect x_t)\Big)dt + \sqrt{2\beta^{-1}} d\vect w_t, 
\end{align}
where $x_t\in\Omega$ denotes the state of the molecular system at time $t$ in state space $\Omega$, $\vect w_t$ denotes standard $n$-dimensional Brownian motion, and $\beta$ the inverse temperature,
i.e.~$\beta = 1/(k_B\mathcal T)$.
Thermostatted Hamiltonain or Langevin dynamics can be treated in the same way as explained herein, so for sake of simplicity we focus on the discussion of diffusive dynamics.
The propagation of probability
densities $\rho=\rho(\vect x,t)$ based on this kind of dynamics in the sense
of $\rho(\vect x,t)dx=\prob[\vect x_t\in [\vect x,\vect x+d\vect x)]$ is governed by
Fokker-Planck equation:
\begin{align}
  \label{eq:disc-fp}
  \frac{\partial \rho}{\partial t} = \ml^\dagger(t) \rho,
\end{align}
where $\ml^\dagger(t)$ is the adjoint of the generator
\begin{align}
  \label{eq:disc-3}
  \ml(t)=\beta^{-1}\Delta_{\vect x}+\Big(-\nabla_x V(\vect x) + E(t)D(\vect x)\Big)\cdot\nabla_{\vect x},
\end{align}
where $\Delta_{\vect x}$ denotes the Laplacian operator and $\nabla_{\vect x}$
the nabla-operator wrt to $\vect x$. 
The periodicity of the external driving force induces the periodicity of the generator,
i.e.~$\ml(t) = \ml(t+T)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial discretization: Master equation}
We will now introduce an appropriate spatial discretization of this kind of non-equilibrium MD --  this is done for reasons of simplicity only; we could completely avoid it for the price of more technical arguments.
For achieving this discretization, we 
introduce a partition of state space $\Omega$ into a finite number of disjoint
sets $\{ \Omega_1, \cdots, \Omega_N\}$ satisfying $\Omega = \cup_i \Omega_i$,
$\Omega_j\cap \Omega_j = \emptyset,\ \forall i\neq j$.
Utilizing the procedure described in Ref.~\cite{latorre2011structure} the original Fokker-Planck equation~(\ref{eq:disc-fp})
is discretized, resulting in a time-inhomogeneous Markov jump process in state
space $S = \{1, \cdots, N\}$ with time-dependent rate
matrix $\vect L(t) \in \mathbb R^{N\times N}$ satisfying
\begin{align}\label{eq:disc-4}
\sum\limits_{j=1}^N L_{ij}(t) & =  0\\ \label{eq:disc-5}
L_{ij}(t) & \ge  0, \quad i\not= j\\
L_{ij}(t) & =  L_{ij}(t+T)
\end{align}
for all real time $t\geq 0$.
Moreover, the rate matrix $L$ has the form $\vect L(t)=\vect L_0+E(t)\vect L_1$
where $E(t)$ is periodic with period $T>0$.
In analogy to \eqref{eq:disc-fp}, the Markov jump process generated by
$\vect L(t)$ transports probability distributions according to the associated Master equation
\begin{align}
  \label{eq:disc-master}
  \frac{d\vect p(t)}{dt} = \vect L^{\top}(t)\cdot \vect p(t)
\end{align}
where $\vect L^{\top}(t)$ denotes the matrix transpose of $\vect L(t)$, $\vect p(t)$ is an $N$-vector denoting the probability distribution on $S$ at time $t$, $p(i,t)$, for example, the probability to be in state $i$ (which corresponds to set $\Omega_i$) at time $t$.
As usual the properties (\ref{eq:disc-4}) and (\ref{eq:disc-5}) of
$\vect L(t)$ guarantee that the total probability mass is conserved,
i.e., if $p(i,0)\ge 0$ componentwise, then $p(i,t)\ge 0$ and $\sum_i
p(i,t) = \sum_ip(i,0)$.
% The solution of the master equation need no be
% periodic.
The temporal evolution of the probability distribution $\vect p(t)$ can be formally written
\begin{align}  \label{eq:disc-8}
\vect p(t)=\myphi(t)\vect p(0)
\end{align}
by using the
associated propagator matrix $\myphi(t)\in\mathbb R^{N\times N}$ that solves
\begin{align}
  \label{eq:disc-master-phi}
  \frac{d}{dt}\myphi(t) = \vect L^{\top}(t)\myphi(t), \quad \myphi(0) = \id .
\end{align}
Since the last equation can be considered column-wise, the propagator matrix inherits column-wise conservation properties:
$\Phi_{ij}(t) \ge  0$
and $\sum\limits_{i=1}^N \Phi_{ij}(t)  =  1$,
that is, $\myphi^{\top}(t)$ is a
stochastic matrix satisfying $\myphi^{\top}(t)\vect e=\vect e$
with $\vect e=(1,\ldots,1)^{\top}\in \mathbb R^N$.
Regarding these considerations, we find
\begin{align}
\label{eq:disc-10}  
\myphi_{ij}(t)=\prob\left(\vect X_t=i\mid \vect X_0=j \right),
\end{align}
where $\vect X_t$ denotes the Markov process generated by $\vect L(t)$.

@@

Since $P$ is a stochastic matrix, its eigenvalues are contained in the unit circle in
the complex plane, i.e., each eigenvalue $\lambda$ (potentially complex-valued)
satisfies $|\lambda|\le 1$. Furthermore $\lambda=1$ is an
eigenvalue with right eigenvector $\vect e=(1,\ldots,1)^{\top}$ and a left eigenvector $\mymu$
satisfying
$\mymu^{\top}P=\mymu^{\top}$.
From now on, we assume $P$ to be irreducible and aperiodic such that the Perron-Frobenius theorem holds, so the eigenvector corresponding to the eigenvalue $\lambda=1$ is  non-negative componentwise, and unique (up to normalization $\sum_i\mu(i)=1$). In this case $\mymu$ is the stationary measure in the sense that
$
 \mymu^{\top} P^m = \mymu^{\top},\ m\in\mathbb N
$,
% \begin{align}
% \label{eq:floq-14}  
% \end{align}
and (more precisely) the asymptotic evolution of an initial probability distribution $\vect p(t=0)$ by the process satisfies
$
\vect p^{\top}(0)\,P^m \to \mymu^{\top},\ m\to\infty,
$
% \begin{align}
% \label{eq:floq-15}  
% \end{align}
so that $\mymu$ can be seen as the quasi-stationary distribution of the non-stationary process.

@@

To illustrate how the discretization works in practice we take the alanine dipeptide system under an oscillatory EF,
as an example, the NEMD simulation of which was
extensively studied in Ref.~\cite{wang2014exploring}.
The system was simulated in a $2.7\times 2.7\times 2.7$~$\textrm{nm}^3$ periodic simulation
region, with one alanine dipeptide molecules described by the CHARMM27 force field~\cite{foloppe2000all} dissolved in 641 TIP3P water molecules~\cite{jorgensen1983comparison}.
The grid-based energy correction map (CMAP)~\cite{mackerell2004extending} was used
to correct the backbone dihedral angle energies.
All simulations were performed by a home-modified
Gromacs~4.6.5~\cite{pronk2013gromacs} with CHARMM27 force field implemented~\cite{bjelkmar2010implementation}.
The alanine dipeptide was put into the local thermostating
environment, with a spherical dynamical region of radius~1.0~nm
centered at the alpha-carbon.
The Langevin thermostat with target temperature $\mathcal T = 300$~K
and timescale $\tau_T = 0.1$~ps was
coupled to the thermostated region.
The whole system was coupled to the Parrinello-Rahman barostat~\cite{parrinello1981polymorphic} (in standard Gromacs implementation) with $\tau_P = 2.0$~ps to
keep the system at 1~Bar. The non-equilibrium trajectories
were integrated by the Leap-frog scheme with a time-step of 0.002~ps.
The short-range van der Waals interactions were cut-off at 1.00~nm, and were smoothed from
0.95~nm to 1.00~nm by the ``\texttt{shift}'' method provide by Gromacs.
The energy conserving 
Particle Mesh Ewald (PME)~\cite{darden1993pme, essmann1995spm} method (``\texttt{pme-switch}'') was
used to compute the long-range electrostatic interaction,
with the same real-space cut-off radius as the van der Waals interactions.
The Gromacs default Fourier spacing of 0.12~nm and B-spline interpolation order of 4 were adopted.
The splitting
parameter was optimized with respect to the electrostatic force computing
accuracy by Gromacs tool \texttt{g\char`_pme\char`_error}~\cite{wang2010optimizing}.
The neighbor list was updated every 5 time-steps with a list-building radius 1.20~nm.
All hydrogen involving covalent bonds were constrained by the LINCS algorithm~\cite{hess1997lincs}, except the water molecules that were constrained by the SETTLE algorithm~\cite{miyamoto2004settle}.
The whole system was driven by a periodic electric field
$E(t) = E_0\sin(2\pi t/T)$ and $D(x) = (1,0,0)^{\top}$
with intensity of the field being $E_0 = 1.0$~V/nm and period being
$T = 10$~ps.
The 20,000 branching trajectories were simulated from 20,000
initial configurations that sample the equilibrium distribution.
% \redc{Write a lot of details on NEMD.}
The equilibrium configurations were prepared by an equilibrium MD simulation
of length $10^6$~ps, along which snapshots were saved every 50~ps.
% In the
% equilibrium simulation the system was coupled to a global Langevin thermostat
% with coupling timescale $\tau_T = 0.5$~ps.
The branching NEMD
trajectories were each 4,000~ps long, and the system reached
non-equilibrium quasi-stationary state in roughly 300~ps.

For this periodically driven molecular system we will first show how to choose an appropriately fine spatial discretization. After validating this discretization we will consider
the time-discretized dynamics generated by the Floquet transition matrix $\vect P$
in comparison to the original NEMD simulation. Finally we will coarse grain this description further by construction of a $3$ state Markov State Model that is able to describe the long-term kinetics of the system correctly.

@@

An important check is the consistency between the
stationary probability density of $\Phi(T)$ (i.e.~the leading eigenvector $\mymu$) and that 
estimated form the original NEMD simulation,
% along which only integral periods $mT$ are considered:
\begin{align}
  \label{eq:num-tmp1}
  \rho_{\textrm{st}}(\phi,\psi) = \lim_{m\rightarrow\infty} \rho (\phi,\psi,mT),
\end{align}
On each NEMD branching trajectory the initial 320~ps are discarded and
the rest of the trajectory in time interval $[320,4000]$~ps is averaged to estimate
the quasi-stationary probability distribution $\rho_{\textrm{st}}$. 
$P$ is computed as described above, and then $\mymu$ is computed as its leading eigenvector. In order to make it comparable to the free energy in the equilibrium case, we take
the logarithm of the distributions, i.e.~$F_{\textrm{st}}(\phi,\psi)=
-k_B\mathcal T\log \rho_{\textrm{st}}(\phi,\psi)$
for NEMD and $F_{\textrm{st}}(\phi,\psi)=
-k_B\mathcal T\log \mu(\phi,\psi)$ for $P$, where $k_B$ is the
Boltzmann constant and $\mathcal T$ is the temperature of the system.
% Under the aformentioned discretization, the stationary distribution is
% defined for the each bin of the dihedral angle space, i.e.~$F_{pq} =
% \int_{ph}^{(p+1)h}d\phi\int_{qh}^{(q+1)h}d\psi
% F_{\textrm{st}}(\phi,\psi), \ 0\leq p,q<N_\dih-1$, where $h = 360/N_\dih$ being
% the size of the bin, $p$ and $q$ here being the bin indexes.
The
results are  compared in Fig.~\ref{fig:num-1}. A good consistency between
the  NEMD simulation and 
$P$ is observed.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Forward and backward committors}
\label{sec:alanine-committor}
\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-msm-1.eps}\\
  \vskip -.7cm
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-msm-1.eps}\\
  \vskip -.7cm
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-msm-1.eps}
  \vskip -.5cm
  \caption{Forward $q^\fwd_{\confaa}$ and backward 
    $q^\bwd_{\confaa}$ committors
    and their difference $q^\fwd_{\confaa} - q^\bwd_{\confaa}$
    computed from the NEMD trajectories (left
    column) and the  discretized dynamics (right column).}
  \label{fig:num-3}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-msm-2.eps}\\
  \vskip -.7cm
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-msm-2.eps}\\
  \vskip -.7cm
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-msm-2.eps}\\
  \vskip -.5cm
  \caption{Forward $q^\fwd_{\beta}$ and backward
    $q^\bwd_{\beta}$ committors
    and their difference  $q^\fwd_{\beta} - q^\bwd_{\beta}$ 
    computed from NEMD trajectories (left
    column) and the  discretized dynamics (right column).}
  \label{fig:num-4}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-3.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-msm-3.eps}\\
  \vskip -.7cm
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-3.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-msm-3.eps}\\
  \vskip -.7cm
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-3.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-msm-3.eps}\\
  \vskip -.5cm
  \caption{Forward $q^\fwd_{\confc}$ and backward $q^\bwd_{\confc}$
    committors
    and their difference  $q^\fwd_{\confc} - q^\bwd_{\confc}$ 
    computed from NEMD trajectories (left
    column) and the  discretized dynamics (right column).}
  \label{fig:num-5}
\end{figure}
